HyperLogLog: Exponentially Bad in Adversarial Settings

### 2020-04-29 docs/HLL_exp_bad/main.tex
more tweaks in expt section

Since the assumptions of S1 appear to us to be unrealistic and since an attacker can escalate from S3 to S4, we focus on the scenarios S2 and S4 in what follows. We first present an attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln} (in their equivalent scenario M1). We then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.

@@

We assumed so far that there are already values coming from other honest users' data in the HLL. We now remove this assumption. The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality estimate of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts. However, the adversary in S4 is able to detect that all buckets are empty and adapt its attack accordingly. The adversary can now target and fill one bucket only. This requires fixing $n+1$ bits of $h(x)$ for each input $x$, allowing the adversary to ``hide'' $2^{t-n-1}$ elements while keeping the HLL cardinality estimate to 1. Other trade-offs between work and the cardinality estimate are of course possible.

@@

\subsection{Experimental Results}
Many different implementations of HLL are available online. Among the repositories collecting more than 350 stars on GitHub, we can cite \cite{clahll} and \cite{datasketch}. HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery \cite{bigquery}, Facebook's Airlift \cite{airlift} and several products of Apache such as Spark \cite{spahll} and Druid \cite{druhll}. It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function. Thus they are vulnerable to our attacks in scenarios S2 and S4.

As a proof-of-concept, we implemented our attacks against the HLL implementation from \cite{clahll}. The attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\textit{\url{https://github.com/PizzaWhisperer/HLLVuln}}}. For simplicity, our items are 4-character long ASCII strings, giving us up to 7,311,616 distinct items that can be inserted. We choose $h$ to be the 32-bit Murmur3 hash function from \cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3  for our attack and consider it reasonable that the adversary should know the hash function in use).
% \kenny{Why do we get to choose $h$? Isn't it fixed by the specific implementation we are attacking?} \mathilde{The library has a place holder for a 32 bit hash function but lets the user choose it. This was one of my concerns and I was wondering if it would make sense to attack \cite{datasketch} instead (hardcoded $h$, but when we discussed it, you said that due to Kerckhoff's principle, this hash function should be public so it does not matter that much.}
We create an HLL sketch using $2^n = 256$ buckets and initialise it with 1,000 random items representing honest users' data.

We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  7,311,616 possible inputs while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario. As a second experiment, and in order to get results that can be directly compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it. Finally, for completeness, we also run the attacks against HLL sketches that are initially empty. All attacks were run 30 times and the outcomes averaged. Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.

These empirical results confirm our impact analysis on several points. First, when attacking a sketch already containing some items, we can see that the adversary is able to hide approximately half of the items in our starting set under S2, and around 83\% under S4. %( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$. Secondly, when attacking an empty sketch, the final cardinality estimate is always exactly 367 $=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by equation (29) in~\cite{hll}, or 1, in respectively S2, S4.

To assess the efficiency of our attacks, we can compare our results from Table~\ref{table:tab2} with the attack presented in Section 5.1 of~\cite{hllvuln}. In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch. Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780``~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality. In comparison, the adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367, providing a reduction by a factor of 340 instead of 5. \kenny{I think we can say more here -- isn't it true that the attack in Section 5.1 of~\cite{hllvuln} is in the M2 scenario? Which is our S1? Whereas the attack you describe here is more like an S2 attack, a much weaker attack model?}

>>>>

Since the assumptions of S1 appear to us to be unrealistic and since an attacker can escalate from S3 to S4, we focus on scenarios S2 and S4 in what follows. We first present an attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln} (in their equivalent scenario M1). We then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.

@@

We assumed so far that there are already values coming from other honest users' data in the HLL. We now remove this assumption and instead assume that the HLL sketch is empty at the beginning of our attacks. The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a worst-case final cardinality estimate of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts. However, the adversary in S4 is able to detect that all buckets are empty and can adapt its attack accordingly. The adversary can now target and fill a single bucket only. This requires fixing $n+1$ bits of $h(x)$ for each input $x$, allowing the adversary to ``hide'' $2^{t-n-1}$ elements while keeping the HLL cardinality estimate to 1. Other trade-offs between work and the cardinality estimate are of course possible in scenario S4.

@@

\subsection{Experimental Results}
Many different implementations of HLL are available online. Among the repositories collecting more than 350 stars on GitHub, we can cite~\cite{clahll} and~\cite{datasketch}. HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery~\cite{bigquery}, Facebook's Airlift~\cite{airlift} and several products of Apache such as Spark~\cite{spahll} and Druid~\cite{druhll}. It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function. Thus they are vulnerable to our attacks in scenarios S2 and S4.

As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}. The attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\textit{\url{https://github.com/PizzaWhisperer/HLLVuln}}}. For simplicity, our items are 4-character long ASCII strings, giving us up to $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks). We choose $h$ to be the 32-bit Murmur3 hash function from \cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
% \kenny{Why do we get to choose $h$? Isn't it fixed by the specific implementation we are attacking?} \mathilde{The library has a place holder for a 32 bit hash function but lets the user choose it. This was one of my concerns and I was wondering if it would make sense to attack \cite{datasketch} instead (hardcoded $h$, but when we discussed it, you said that due to Kerckhoff's principle, this hash function should be public so it does not matter that much.}
We create an HLL sketch using $2^n = 256$ buckets and initialise it with 1,000 random items representing honest users' data.

We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $7,311,616$ possible inputs while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario. As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it. Finally, for completeness, we also run the attacks against HLL sketches that are initially empty. All attacks were run 30 times and the outcomes averaged. Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.

These empirical results confirm our impact analysis on several points.

First, when attacking a sketch already containing some items, we can see that the adversary is on average able to hide about half of the items in our starting set under S2 (as predicted by our theoretical analysis), and about 83\% under S4. %( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$, so the results are also consistent with our theoretical analysis.

Secondly, when attacking an empty HLL sketch in S2, the final cardinality estimate is always exactly $367=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by equation (29) in~\cite{hll}. For S4, the final cardinality estimate is always exactly 1 (this comes from rounding up the estimated cardinality of $\alpha = 0.72$ that arises from our attack targeting a single bucket). Note that in S4 with an initially empty HLL sketch, we insert (on average) 14,220 items into the sketch from our candidate set of $7,311,616$ items; this is again consistent with our theoretical predictions of $2^{t-n-1} = 2^{22.8-9} = 2^{13.8} = 14,263$. Of course, we could insert many more items into an initially empty HLL sketch in S4 by relaxing the single bucket restriction. This would come at the price of allowing the cardinality estimate to increase beyond 1.

We can also compare our results from Table~\ref{table:tab2} with results from the attack presented in Section 5.1 of~\cite{hllvuln}. In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch. Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780''~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality. The authors present this attack as being in their M2 (our S1) scenario, without providing a detailed explanation. By looking at the Redis open-source code, we were able to obtain full details of the implementation that any attacker could obtain, and thus argue that based on this knowledge, the adversarial scenario is actually M1 (S2 in this work). This makes the comparison between the results of~\cite{hllvuln} and our S2 results appropriate. The S2 adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367. Compared to~\cite{hllvuln}, we can insert twice as many items and reduce the cardinality estimate by a factor of 340 instead of 5.
